Multimodal Information Extraction of Supermarket Leaflets

### 2025-02-21 paper/main_ilyi.tex
Add recent changes to paper.

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

@@

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{coling}

@@

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

@@

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}

@@

\author{Xincheng Liao\textsuperscript{}, Junwen Duan\textsuperscript{}\thanks{\ \ Corresponding author. Email: \href{mailto:jwduan@csu.edu.cn}{jwduan@csu.edu.cn}}, Yixi Huang\textsuperscript{}, Jianxin Wang\textsuperscript{} \\
Hunan Provincial Key Lab on Bioinformatics, School of Computer Science and Engineering, \\
Central South University, Changsha, Hunan, China \\
\texttt{\{ostars, jwduan, yx.huang\}@csu.edu.cn, jxwang@mail.csu.edu.cn} \\
\href{https://github.com/OStars/RUIE}{https://github.com/OStars/RUIE}
}

@@

\section{Introduction}
\subsection{Motivation}
\subsection{Problem Statement}
\subsection{Contributions}

@@

@@

\section{Information Extraction of Supermarket Deals}
    % - Subsequent to detection and separation of deals, the goal is to extract useful information from it.
    % - Difficulties
    %     - High variety of layouts, colors, fonts, sizes
    %         - Deal sometimes have superscripted decimals and no separation dot.
    %         - Original prices are sometimes striked through.
    %         - Deal prices are sometimes in a very different font to the ones it was trained on.
    %         - Overlapping elements
    %     - Multimodal information
    %         - Product image: visual
    %         - Text: alphabetical + numerical
    %     - OCR errors
    %     - Eigennamen (mostly brand names)
    %     - German language
    % - Goal: Extract 
    %     - Product name
    %     - brand
    %     - original price
    %     - deal price
    %     - unit
Subsequently to the detection and separation of supermarket deals, the overarching goal of using these in applications requires the extraction of the various useful information from these deals. 
The task of extracting meaningful and structured information from supermarket deal images is a complex and multi-faceted problem, requiring the integration of various modalities, overcoming challenges posed by noisy data, and leveraging deep learning techniques. This section provides a detailed exploration of the underlying challenges, methodologies, and the theoretical and practical framework needed to perform such extraction.

@@

\subsection{Architectural Approaches to Information Extraction}
- Classical methods: Rule-based, template-based
- Traditional methods: Dividing into multi-stage
    - 1. Text detection
    - 2. Text recognition
    - 3. Key-value pair extraction
- OCR + LLM
- End2End: Vision Encoder Decoder Model
    - Donut 
- E2E: LVLM

\subsection{Dataset Creation}
Since there was no dataset to evaluate or train, a custom dataset was created by manually annotating a collection of supermarket deal images. The dataset includes images with diverse layouts, fonts, and colors to ensure robustness. Each image is annotated with bounding boxes for text regions and corresponding labels for product name, brand, original price, deal price, and unit. The dataset is split into training, validation, and test sets to facilitate model development and evaluation. 
% Table \ref{tab:dataset} provides an overview of the custom dataset and Figure \ref{fig:dataset} shows sample images with annotations.
% TODO: Add table and figure

\subsection{Optical Character Recognition}
Optical Character Recognition (OCR) is a critical component in the information extraction pipeline. Several OCR solutions were evaluated to determine the most suitable one for our task. The evaluated OCR tools include Tesseract, EasyOCR, PaddleOCR, and DocTR. Each tool was assessed based on its accuracy, speed, and ability to handle diverse fonts and layouts.

\subsection{Experiments}
\indent{\textbf{Setup}.}
The experiments were conducted to compare the performance of different OCR solutions and end-to-end models for information extraction. The evaluation metrics include precision, recall, and F1-score for each entity type (product name, brand, original price, deal price, and unit). The experiments were performed on the test set of the custom dataset.

\indent{\textbf{Results}.}
The results of the experiments indicate that the end-to-end models, such as the Vision Encoder Decoder Model (e.g., Donut) and Large Vision-Language Models (LVLM), outperform traditional OCR-based methods in terms of accuracy and robustness. The OCR + LLM approach also shows promising results, particularly in handling OCR errors and extracting key-value pairs.

\subsection{Evaluation}
The evaluation of the models was conducted using standard metrics, including precision, recall, and F1-score. The comparison between LVLM, OCR + LLM, and Donut models highlights the strengths and weaknesses of each approach. The results demonstrate that end-to-end models provide superior performance in extracting structured information from supermarket deal images.

\section{Application}
\subsection{Database}

\subsection{Optical Character Recognition}

    \subsection{Experiments}
    \indent{\textbf{Setup}.}
    - Common OCR solutions: Tesseract, EasyOCR, PaddleOCR, DocTR
    - Goal: Which OCR solution is the best for our task?
    -

    \subsection{Evaluation}
    - LVLM, OCR+LLM, Donut
    - OCR + LLM Vergleich

>>>>

% !TEX root = main_ilyi.tex

% for pdfLaTeX
\pdfoutput=1

@@

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage{lib}

@@

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}

@@

\usepackage{microtype}

\usepackage{inconsolata}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}

@@

\author{Ilyesse Hettenbach\textsuperscript{}, Gabriel Schurr\textsuperscript{} \\
HKA, Karlsruhe, Germany \\
\texttt{\{heil1012, scga1011\}@h-ka.de} \\
\href{https://github.com/ilyii/leaflets}{https://github.com/ilyii/leaflets}
}

@@

\section{Introduction}

\subsection{Motivation}
If one thinks of a paper-heavy country, Germany is likely to come to mind. The country's affinity for printed materials is deeply ingrained in its culture, with activities such as browsing through supermarket leaflets serving as a familiar ritual for many individuals. These leaflets have long been an essential medium for consumers to plan their purchases, discover special offers, and compare prices. The tactile nature of printed leaflets, the excitement of spotting discounts, and the satisfaction of making a well-informed shopping decision contribute to their enduring appeal.
However, the rapid advancement of digital technology is transforming consumer behavior. While some individuals still find browsing through physical leaflets a nostalgic and even relaxing experience, younger generations are increasingly turning to digital alternatives for their shopping needs. The modern shopper expects convenience, efficiency, and instant access to information, yet the digital transformation of supermarket leaflets lags significantly behind other aspects of e-commerce and retail technology.

Despite the existence of applications that compile digital versions of these leaflets, their usability and overall quality remain suboptimal. Many digital leaflets are plain documents, failing to leverage the full potential of interactivity, searchability, and intelligent recommendation systems. There is a pressing need for sophisticated digital solutions that not only present supermarket deals but also enable consumers to interact with them in a meaningful way.

Possibly, one can imagine a seamless digital platform where users can search for specific items across multiple supermarket chains, apply dietary filters to instantly highlight suitable products, or receive personalized recommendations based on past preferences. A tool that allows consumers to create a dynamic shopping list, track price trends over time, and receive real-time updates on the best deals would revolutionize the way people approach grocery shopping. Yet, these features remain largely absent from existing digital leaflet solutions. Furthermore, while price comparison websites and applications have become commonplace for electronics, fashion, and travel, the grocery sector remains an overlooked frontier. Consumers are often left navigating multiple supermarket websites or manually cross-referencing prices to find the best deals—a tedious and inefficient process. A comprehensive digital ecosystem that integrates grocery price comparisons, personalized discounts, and AI-driven shopping assistants could bridge this gap and act as a workaround, offering an enhanced, data-driven shopping experience tailored to individual needs.

The future of supermarket deal discovery should not be a static PDF or an unorganized collection of scanned pages—it should be a dynamic, interactive, and intelligent experience that aligns with the evolving digital landscape.

\subsection{Problem Statement}
The primary objective of this research is to develop an intelligent system for extracting, structuring, and utilizing information from supermarket leaflets in a digital format. Given a set of supermarket leaflets \(\mathcal{L} = \{L_1, L_2, ..., L_n\}\), each containing a collection of product deals with specific attributes such as product name, brand, price, discount, product image and other product specifications, the goal is to extract and normalize these attributes to enable efficient querying and comparison across multiple supermarket chains.

Each leaflet \( L_i \) consists of a set of $m$ product entries \( P_i = \{p_1, p_2, ..., p_m\} \), where each product \( p_j \) is associated with a set of varying attributes:

\begin{equation}
    p_j = \{a_1, a_2, ..., a_k\},   
\end{equation}
where each attribute \( a_k \) represents a specific entity. Common attributes include:
\begin{itemize}
    \item Product Name
    \item Brand
    \item Original Price
    \item Deal Price
    \item Unit (e.g., weight, volume)
    \item Product Image
\end{itemize}
among others. Each attribute may or may not be present in a given product entry, and the format and layout of these attributes can vary significantly across different leaflets.

To enable advanced applications such as the abovementioned, one has to tackle different challenges, such as:
% Nummerierung der Challenges
\begin{enumerate}
    \item Detecting and separating deals $p$ from the leaflet $L$.
    \item Extracting structured information from the detected deals.
    \item Normalizing the extracted information to ensure consistency and comparability.
    \item Validating the extracted information to ensure accuracy.
\end{enumerate}

Therefore, a function $$f: \mathcal{L} \to \mathcal{D}$$ has to be designed, where $\mathcal{D}$ represents the structured information extracted from the leaflets. The function $f$ should minimize the extraction error $E$.

\subsection{Contributions}  
This work presents a comprehensive approach to digitalizing and structuring supermarket leaflet information by leveraging advanced computer vision, OCR, and language models. The key contributions of this research are as follows:

% 1. **Development of a Detection and Extraction Pipeline for Supermarket Leaflets**  
\indent{\textbf{Development of a Detection and Extraction Pipeline for Supermarket Leaflets.}} We propose a modular and efficient pipeline for detecting, extracting, and structuring information from supermarket leaflets. This pipeline integrates OCR-based text extraction, image processing, and layout understanding to accurately associate textual and visual elements such as product names, prices, descriptions, and images. Additionally, we incorporate normalization techniques to standardize extracted data across different retailers and formats.

% 2. **Design of an Interactive Application for Deal Browsing and Price Comparison**  
\indent{\textbf{Design of an Interactive Application for Deal Browsing and Price Comparison.}} To facilitate practical use, we wrap the extracted data into a user-friendly application that enables consumers to browse, filter, and compare deals across multiple supermarket chains. This application includes basic functionalities to search, track and compare deals across different supermarket chains.

% 3. **Comprehensive Evaluation of Information Extraction Models**  
\indent{\textbf{Comprehensive Evaluation of Information Extraction Models.}} We systematically evaluate multiple state-of-the-art models for supermarket leaflet information extraction, including hybrid OCR+LLM approaches, end-to-end Vision-Language Models (VLMs), and Large Vision-Language Models (LVLMs). Our evaluation covers accuracy metrics such as per-entity correctness, Levenshtein distance, and robustness across different supermarket layouts and textual formats. This analysis provides insights into the trade-offs between computational efficiency and extraction performance.

% 4. **Training of Custom Deal Detection and Information Extraction Models**  
\indent{\textbf{Training of Custom Deal Detection and Information Extraction Models.}} To further improve extraction performance, we develop and fine-tune specialized models for deal detection and information extraction. This includes training custom object detection models to identify product regions within leaflets, as well as domain-specific sequence-to-sequence models for structured text extraction. These models are designed to enhance recognition accuracy, particularly for complex layouts and noisy scan conditions.

% 5. **Creation of a Benchmark Dataset: Leaflet-IE**  
\indent{\textbf{Creation of a Benchmark Dataset: Leaflet-IE.}} We introduce **Leaflet-IE**, a novel dataset curated for supermarket leaflet information extraction. The dataset consists of annotated images spanning multiple supermarket chains, with labeled entities including product names, prices, discounts, brands, and categories. This dataset serves as a benchmark for future research in structured information extraction from promotional materials and contributes to advancing OCR and multimodal learning for real-world retail applications.

@@

\section{Nature of Supermarket Leaflet Data}
Supermarket leaflets constitute a complex and heterogeneous data source, characterized by multimodal content that combines structured and unstructured information. This data presents unique challenges for computational processing due to its inherent ambiguities, multi-instance object representations, and spatial dependencies. The reader is encouraged to simultaneously inspect the visual representation of these challenges in \figref{fig:leaflet_data} while proceeding with the discussion.

One of the primary challenges lies in the ambiguity of price recognition, where a single product may be associated with multiple price points, such as discounted prices, bulk purchase offers, or tiered pricing structures. This ambiguity is further exacerbated by errors in punctuation extraction, particularly in recognizing decimal places, which often arise from font-specific artifacts or noisy backgrounds. Resolving such issues requires advanced plausibility checks, contextual reasoning, and robust numerical interpretation mechanisms to ensure accurate data extraction. Additionally, the non-strict one-to-one mapping between textual and visual elements introduces further complexity. For instance, a single product image may represent multiple items, or textual descriptions may not align precisely with their corresponding visual representations. This misalignment necessitates sophisticated multimodal fusion techniques to bridge the gap between visual and textual data, ensuring accurate association and interpretation of product information.

The inconsistent layouts of supermarket leaflets further complicate data extraction. Promotional deals and product information are often placed in varying positions across different leaflets, making it difficult to design a universal extraction method that generalizes across diverse formats. Text elements frequently overlap with product images or other textual information, leading to misidentification or incorrect parsing of data. These challenges are compounded by the variability in fonts, sizes, and weights used within a single leaflet, which complicates the differentiation between product names, prices, and promotional tags. The presence of multilingual content and special characters adds another layer of complexity, as leaflets may contain text in multiple languages or stylized symbols that require optical character recognition (OCR) and natural language processing (NLP) models capable of handling diverse linguistic structures and alphabet systems.

Promotion labels and discount representations introduce additional challenges due to their non-standard or highly stylized text formats. Identifying the exact nature of promotions, such as "Buy 1, get 1 free" or "20\% off," requires precise extraction and interpretation, particularly when the text is embedded in noisy backgrounds or overlaps with other elements. Furthermore, the temporal context of deals and promotions adds a critical dimension to the extraction process. Some offers may only be valid for specific periods, and extracting and associating the correct date ranges with these offers can be challenging, especially when the information is not explicitly stated or is presented in ambiguous formats.

Therefore, the nature of supermarket leaflet data necessitates an advanced toolkit of visual and textual processing, analysis, extraction and generation techniques to effectively interpret and utilize the information contained within these documents. The subsequent sections will delve into the methodologies and models used to address these challenges and extract structured information from supermarket leaflets.

@@

\section{Information Extraction of Supermarket Deals}
Subsequently to the detection and separation of supermarket deals, the overarching goal of using these in applications requires the extraction of the various useful information from these deals. 
The task of extracting meaningful and structured information from supermarket deal images is a complex and multi-faceted problem, requiring the integration of various modalities, overcoming challenges posed by noisy data, and leveraging deep learning techniques. This section provides a detailed exploration of the underlying challenges, methodologies, and the theoretical and practical framework needed to perform such extraction.

@@

\subsection{Architectural Approaches to Information Extraction}
Among the various existing architectural approaches to information extraction, several methodologies have been developed to address the inherent complexities in supermarket deal images. These methods can be categorized into classical approaches, multi-stage traditional models, hybrid OCR-LLM systems, and end-to-end deep learning frameworks. Each paradigm offers distinct advantages, and their applicability depends on computational resources, dataset characteristics, and performance requirements.

\subsubsection{Traditional and Hybrid Approaches for Supermarket Leaflet Information Extraction}  
Information extraction from supermarket leaflets has traditionally been approached through a pipeline of sequential sub-tasks, each addressed by specialized systems. These methods, while extensively studied and methodologically diverse, are typically divided into three main stages: text detection, text recognition, and key-value pair extraction. Although effective in certain scenarios, these approaches suffer from inherent limitations, including computational overhead, error propagation, and suboptimal performance on edge cases. Recent advancements in large language models (LLMs) have led to the development of hybrid OCR-LLM architectures, which combine the precision of optical character recognition (OCR) with the contextual reasoning capabilities of LLMs. Below, we formalize these approaches and discuss their strengths and limitations.

Traditional architectures decompose the problem into independent sub-tasks, each handled by specialized models. While this modularity allows for targeted optimization, it introduces complexity and inefficiencies due to the lack of end-to-end training and error propagation across stages.

The first stage, \emph{text detection}, involves identifying regions in the image \( I \in \mathbb{R}^{H \times W \times 3} \) that contain textual information. This is typically achieved using rule-based systems or neural networks designed for small object detection. The output is a set of \( M \) bounding boxes \( \{b_i\}_{i=1}^{M} \), where each \( b_i = (x_{\text{min}}, y_{\text{min}}, x_{\text{max}}, y_{\text{max}}}) \) defines the coordinates of a text region. Modern approaches often employ convolutional neural networks (CNNs) or transformer-based architectures to improve robustness to varying fonts, orientations, and overlapping text.

Each detected bounding box \( b_i \) is cropped and processed by a \emph{text recognition} model to convert the visual representation into a textual one. This task is challenging due to decoding errors caused by noisy backgrounds, stylized fonts, and ambiguous characters. Commonly used models include convolutional recurrent neural networks (CRNNs) \cite{shi2016crnn} and transformer-based recognizers. Given the set of bounding boxes \( \{b_i\}_{i=1}^{M} \), the goal is to predict the corresponding textual entities \( \{t_i\}_{i=1}^{M} \):  
\[
\{t_i\}_{i=1}^{M} = \arg\max_{\{t_i\}_{i=1}^{M}} P(\{t_i\}_{i=1}^{M} | \{b_i\}_{i=1}^{M}).
\]  
The text recognition stage is often optimized using connectionist temporal classification (CTC) loss or sequence-to-sequence objectives, depending on the model architecture.

The final stage structures the extracted text into meaningful entities, such as product names, prices, and promotional tags. Named entity recognition (NER) models or graph neural networks (GNNs) are commonly used to learn associations between textual elements. The objective is to maximize the alignment score \( \Psi(v_i, v_j) \) between related entities:  
\[
\mathcal{A}^* = \arg\max_{\mathcal{A} \subseteq E} \sum_{(i,j) \in \mathcal{A}} \Psi(v_i, v_j),
\]  
where \( E \) represents textual adjacency relationships. This stage is critical for resolving ambiguities, such as associating prices with the correct products.

While traditional pipelines are widely used, they suffer from several limitations. The need for multiple independently trained models increases computational overhead and complicates system management. Additionally, the lack of end-to-end training often leads to suboptimal performance, particularly on edge cases or unseen data, as errors propagate through the pipeline.

The rise of large language models (LLMs) has enabled the development of hybrid OCR-LLM architectures, which combine the strengths of OCR tools with the linguistic and reasoning capabilities of LLMs. These approaches retain the traditional OCR pipeline for converting visual text into machine-readable form but leverage LLMs to enhance robustness through contextual reasoning and structured output generation.

In a hybrid OCR-LLM system, the OCR stage processes the input image \( I \) to produce a set of unstructured text fragments \( T = \{t_i\}_{i=1}^{M} \) with associated spatial coordinates. These fragments are serialized into a prompt \( \mathbf{P} \) and fed into a pretrained LLM \( F_\theta \), which performs contextual parsing and generates structured output \( \hat{y} \):  
\[
\hat{y} = F_\theta (T).
\]  
The LLM leverages its pretrained knowledge to resolve ambiguities, infer relationships between text fragments, and generate outputs in a structured format (e.g., JSON). For example, it can distinguish between bulk pricing (e.g., "2 for \$4") and unit pricing (e.g., "\$2 each") through arithmetic reasoning embedded in its attention mechanisms.

Hybrid OCR-LLM models offer several advantages over traditional pipelines. By integrating contextual reasoning, they reduce error propagation and improve robustness to noisy or ambiguous inputs. Additionally, the use of pretrained LLMs eliminates the need for task-specific training, simplifying system development. However, these models still rely on the accuracy of the OCR stage, and misrecognitions (e.g., "O" vs. "0") can propagate to the LLM. Furthermore, the computational cost of LLM inference can be prohibitive for large-scale applications.

\subsubsection{End-to-End Models}
End-to-End (E2E) approaches eliminate pipeline fragmentation by learning a unified mapping $F: I \to \mathcal{D}$.

\indent{\textbf{Vision-Encoder-Decoder Models.}} These models represent a paradigm shift in document understanding by unifying text detection, recognition, and semantic parsing into a single end-to-end trainable framework. Among these, the Donut architecture \cite{kim2022donut} has emerged as a state-of-the-art model for tasks such as information extraction from supermarket leaflets. Donut eliminates the need for intermediate OCR steps by directly mapping input images to structured outputs, leveraging a transformer-based encoder-decoder architecture. Below, we formalize the Donut model, its mathematical foundations, and its application to supermarket leaflet data.

Donut consists of two primary components: a vision encoder and a text decoder, both built on transformer layers. Given an input image \( I \in \mathbb{R}^{H \times W \times 3} \), the model generates a structured output \( \mathbf{Y} \) in an autoregressive manner. The encoder extracts high-level visual features, while the decoder translates these features into a sequence of tokens representing the desired output format.

The \emph{vision encoder} processes the input image \( I \) into a sequence of visual embeddings \( \mathbf{Z} = \{z_i\}_{i=1}^{L} \), where \( L \) is the sequence length. This is achieved using a Swin Transformer \cite{liu2021swin}, which partitions the image into non-overlapping patches and applies hierarchical self-attention to capture both local and global context. Let \( \mathbf{X} \in \mathbb{R}^{N \times (P^2 \cdot C)} \) denote the patch embeddings, where \( N \) is the number of patches, \( P \) is the patch size, and \( C \) is the number of input channels. The Swin Transformer computes multi-head self-attention (MSA) and feed-forward network (FFN) operations as follows:  
\[
\mathbf{X}' = \text{MSA}(\text{LayerNorm}(\mathbf{X})) + \mathbf{X},  
\]  
\[
\mathbf{Z} = \text{FFN}(\text{LayerNorm}(\mathbf{X}')) + \mathbf{X}',
\]  
where LayerNorm denotes layer normalization. The output \( \mathbf{Z} \) serves as the visual representation for the decoder.

The \emph{text decoder} is a transformer-based autoregressive model that generates the output sequence \( \mathbf{Y} = (y_1, y_2, \dots, y_T) \) conditioned on the visual embeddings \( \mathbf{Z} \). At each time step \( t \), the decoder predicts the next token \( y_t \) based on the previously generated tokens \( y_{<t} \) and the encoder output \( \mathbf{Z} \). The decoder computes self-attention over the output sequence and cross-attention over the visual embeddings:  
\[
\mathbf{Q} = \mathbf{W}_Q \mathbf{H}_{t-1}, \quad \mathbf{K} = \mathbf{W}_K \mathbf{Z}, \quad \mathbf{V} = \mathbf{W}_V \mathbf{Z},
\]  
\[
\text{Cross-Attention}(\mathbf{Q}, \mathbf{K}, \mathbf{V}) = \text{softmax}\left(\frac{\mathbf{Q} \mathbf{K}^\top}{\sqrt{d_k}}\right) \mathbf{V},
\]  
where \( \mathbf{H}_{t-1} \) is the hidden state at step \( t-1 \), and \( \mathbf{W}_Q, \mathbf{W}_K, \mathbf{W}_V \) are learnable projection matrices. The final token probabilities are computed as:  
\[
P(y_t | y_{<t}, \mathbf{Z}) = \text{softmax}(\mathbf{W}_o \mathbf{H}_t),
\]  
where \( \mathbf{W}_o \) is an output projection matrix. The model is trained end-to-end using a cross-entropy loss:  
\[
\mathcal{L} = -\sum_{t=1}^T \log P(y_t^* | y_{<t}, \mathbf{Z}),
\]  
where \( y_t^* \) is the ground-truth token at step \( t \).

Donut is particularly well-suited for supermarket leaflet data due to its ability to handle multimodal inputs and generate structured outputs without intermediate OCR steps. The model can directly process leaflet images and produce JSON representations of product information. By eliminating the need for OCR, it avoids error propagation and simplifies the pipeline. Its end-to-end trainability ensures optimal performance, and its ability to generate structured outputs reduces the need for post-processing. However, Donut requires large-scale pretraining on diverse document datasets, which can be computationally expensive. Additionally, its performance on low-resolution or highly noisy images may be suboptimal, as the vision encoder relies on high-quality visual features.

\indent\textbf{Large Vision-Language Models.} Large Vision-Language Models (LVLMs) represent the latest advancement in multimodal learning, combining the strengths of LLMs and vision encoders to achieve state-of-the-art performance on a wide range of tasks. These models, such as CLIP \cite{radford2021clip} and ALIGN \cite{li2022align}, leverage large-scale pretraining on diverse multimodal datasets to learn robust representations of visual and textual information. Below, we formalize the ALIGN model and discuss its application to supermarket leaflet data.

\section{Large Vision-Language Models for Supermarket Leaflet Understanding}  
Large Vision-Language Models (LVLMs) represent a transformative advancement in multimodal artificial intelligence, unifying visual perception and linguistic reasoning within a single cohesive framework. These models, such as LLaVA \cite{liu2023llava} and GPT-4V \cite{openai2023gpt4}, eliminate the traditional separation between vision and language processing by jointly encoding images and text into a shared latent space. For supermarket leaflet analysis, LVLMs enable end-to-end extraction of structured information directly from raw images, bypassing explicit OCR stages while leveraging contextual synergies between visual and textual data. We formalize the architectural principles, mathematical foundations, and application-specific adaptations of LVLMs below.

---

### **Architecture and Mathematical Formulation**  
LVLMs integrate a vision encoder, a language model, and cross-modal fusion mechanisms into a unified architecture. Given an input leaflet image \( I \in \mathbb{R}^{H \times W \times 3} \), the model generates structured textual output \( \mathbf{Y} \) through joint reasoning over visual and linguistic features.

#### **Vision Encoder**  
The vision encoder processes \( I \) into a sequence of patch embeddings \( \mathbf{Z} \in \mathbb{R}^{L \times d_v} \), where \( L \) is the number of patches and \( d_v \) is the embedding dimension. A Vision Transformer (ViT) \cite{dosovitskiy2020vit} is commonly employed, partitioning \( I \) into \( N \) non-overlapping patches \( \{\mathbf{p}_i\}_{i=1}^N \), linearly projecting each patch, and applying transformer layers with self-attention:  
\[
\mathbf{z}_i = \text{MSA}(\mathbf{p}_i) + \mathbf{p}_i, \quad \mathbf{Z} = [\mathbf{z}_1; \mathbf{z}_2; \dots; \mathbf{z}_N],
\]  
where MSA denotes multi-head self-attention. For supermarket leaflets, the encoder captures spatial relationships critical for disambiguating overlapping text and product layouts.

#### **Cross-Modal Fusion**  
Visual embeddings \( \mathbf{Z} \) are aligned with textual representations through a cross-modal projector, which maps \( \mathbf{Z} \) into the language model's embedding space. Let \( \mathbf{W}_p \in \mathbb{R}^{d_v \times d_t} \) denote a learnable projection matrix, where \( d_t \) is the language model's hidden dimension. The projected visual tokens \( \mathbf{V} \) are computed as:  
\[
\mathbf{V} = \mathbf{Z} \mathbf{W}_p.
\]  
These tokens are concatenated with text token embeddings \( \mathbf{E}_{\text{text}} \in \mathbb{R}^{T \times d_t} \) to form a unified input sequence:  
\[
\mathbf{H}_{\text{in}} = [\mathbf{V}; \mathbf{E}_{\text{text}}].
\]

#### **Language Model Decoding**  
The language model (e.g., LLaMA \cite{touvron2023llama}) processes \( \mathbf{H}_{\text{in}} \) autoregressively to generate structured output \( \mathbf{Y} = (y_1, \dots, y_T) \). At each step \( t \), cross-attention layers attend to visual tokens \( \mathbf{V} \) while self-attention layers model dependencies within the output sequence:  
\[
\mathbf{Q}_t = \mathbf{W}_Q \mathbf{h}_{t-1}, \quad \mathbf{K}_v = \mathbf{W}_K \mathbf{V}, \quad \mathbf{V}_v = \mathbf{W}_V \mathbf{V},
\]  
\[
\mathbf{h}_t' = \text{softmax}\left(\frac{\mathbf{Q}_t \mathbf{K}_v^\top}{\sqrt{d_t}}\right) \mathbf{V}_v + \mathbf{h}_{t-1},
\]  
\[
P(y_t | y_{<t}, \mathbf{Z}) = \text{softmax}(\mathbf{W}_o \mathbf{h}_t'),
\]  
where \( \mathbf{W}_Q, \mathbf{W}_K, \mathbf{W}_V, \mathbf{W}_o \) are learnable parameters. The model is trained end-to-end using a cross-entropy loss:  
\[
\mathcal{L} = -\sum_{t=1}^T \log P(y_t^* | y_{<t}, \mathbf{Z}),
\]  
where \( y_t^* \) is the ground-truth token.

---

### **Application to Supermarket Leaflets**  
LVLMs excel at extracting structured product information (e.g., prices, promotions) by jointly reasoning over visual and textual cues. For instance, given a leaflet image with the text "\$2.99" adjacent to an apple image, the model infers the association through cross-attention, generating output:  
\[
\{\text{product: "Apples", price: 2.99, promotion: "Weekly Special"}\}.
\]  
Key advantages include:  
1. **Ambiguity Resolution**: Differentiates prices from non-price numerals (e.g., "2 lbs") via contextual alignment of text and images.  
2. **Layout Robustness**: Leverages spatial attention to associate scattered text fragments (e.g., disclaimers linked to products).  
3. **Multilingual Support**: Inherits multilingual capabilities from pretrained language models, handling leaflets with mixed languages.

---

### **Advantages and Limitations**  
LVLMs eliminate error-prone OCR pipelines and enable zero-shot generalization through their pretrained knowledge. Their unified architecture simplifies deployment and reduces computational overhead compared to hybrid OCR-LLM systems. However, limitations persist:  
- **Training Complexity**: Requires large-scale multimodal datasets (e.g., leaflet images paired with JSON outputs) for fine-tuning.  
- **Hallucination Risks**: May generate plausible but incorrect details if visual grounding is weak (e.g., inventing prices for occluded text).  
- **Resource Intensity**: High memory and compute demands during inference, limiting real-time applications.

---

### **Conclusion**  
Large Vision-Language Models redefine supermarket leaflet information extraction by unifying visual and linguistic reasoning into an end-to-end framework. Their ability to directly map raw images to structured outputs, bypassing intermediate OCR stages, positions them as a promising solution for complex retail document understanding. Future work should focus on enhancing visual grounding, optimizing inference efficiency, and expanding multilingual support to address current limitations.

\subsection{Dataset Creation}
Since the availability of German supermarket leaflet data is literally not findable, a custom dataset was created by manually annotating a collection of supermarket deal images. The dataset includes images from the most common supermarket chains, each with unique layout, fonts, and colors to, in general, ensure the highest diversity and robustness possible. Each image is annotated with a corresponding label file that includes the desired entities as well as the image identificator. The dataset is split into a training (~ 80\%) and a validation set (~ 20\%) for each model to be trained and evaluated on. Table \tabref{tab:ie_dataset} provides an overview of the dataset with the sample size for each entity and \figref{fig:ie_dataset_samples} shows some sample images with their annotations.

\begin{table}[ht]
\centering
\caption{Leaflet-IE Dataset}
\label{tab:ie_dataset}
\begin{tabular}{lc}
\toprule
Entity           & Sample Size \\
\midrule
\code{image\_id}        & 372  \\
\code{brand}            & 357  \\
\code{product\_name}    & 370  \\
\code{original\_price}  & 286  \\
\code{deal\_price}      & 372  \\
\code{weight}           & 369  \\
\bottomrule
\end{tabular}
\end{table}

Even though the Leaflet-IE dataset is relatively small, the reader will be able to see that the dataset is sufficient to evaluate the performance of different IE appraoches as well as to train an competitive end-to-end model. However, as the reader may notice, the dataset is solely focused on single product deals at this point and explicitly does not include

\begin{figure*}[ht]
\centering
\fbox{\parbox[c][3cm][c]{5cm}{\centering Figure Placeholder: Sample Annotation}}
\caption{Sample images with annotations from the custom dataset.}
\label{fig:ie_dataset_samples}
\end{figure*}

\section{Experiments}

The primary objective of these experiments is to identify the most effective approach for extracting structured information from supermarket leaflets. The evaluation considers general performance trends and use cases to determine candidate methodologies, including Hybrid OCR + LLM, end-to-end Vision-Encoder-Decoder models, and LVLMs.
The order of the experiments is as follows:
- Plain Deal OCR Performance
- OCR + LLM Performance
- Donut Fine-Tuning Performance
- MAIN: Comparison of Hybrid OCR+LLM, Donut, and LVLM

\subsection{Implementation Details}
The experiments were conducted on two distinct computational workstations - station 1 equipped with an NVIDIA RTX 4080 GPU and station 2 with an NVIDIA GTX 1080 GPU. Therefore, the experiments could be conducted in parallel, ensuring efficient model training and evaluation.

The software stack includegraphics Python with PyTorch as the primary deep learning framework, HuggingFace Transformers for leveraging pre-trained models, and OpenCV for image preprocessing and data augmentation.
 PyTorch, HuggingFace Transformers, and various other libraries for data pre-and post-processing, visualization, data management and monitoring \footnote{The codebase is available at \href{https://github.com/ilyii/leaflets}{https://github.com/ilyii/leaflets}}.

% SUB-STUDIES
\subsection{Plain Deal OCR Performance}
The aim of this experiment is to evaluate the performance of various OCR models in extracting text from supermarket leaflets. The evaluation focuses on the accuracies of the extracted entities from \tabref{tab:ie_dataset} and the impact of different normalization levels on the recognition performance. The results provide insights into the robustness and accuracy of different OCR models in handling the complexities of supermarket leaflet data.

The most commonly used OCR models include:
\begin{itemize}
    \item Tesseract,
    \item EasyOCR,
    \item PaddleOCR,
    \item DocTR.
\end{itemize}
These models have been trained on a variety of datasets and are known for their robustness and accuracy in text extraction. Since it is likely that none of the OCR models have seen this data before, all 372 samples of the Leaflet-IE dataset were used for evaluation. While OCR alone is not capable of generating key-value mappings, the evaluation used a different set of metrics:

\begin{itemize}
    \item \textbf{Accuracy}: Measures whether the expected value is present in the OCR output.
    \item \textbf{N-gram Accuracy}: Determines if an n-gram substring of the expected value exists in the OCR output, computed for \(n \in \left[\frac{|v|}{2}, |v| - 1\right]\), where \(|v|\) represents the string length of the ground truth entity value.
\end{itemize}
